Classless Java

### 2015-11-14 paper/sections/Introduction.tex
Ok, pdflatex is now working fine. I submit a couple of typo fix to introduction/abstract and some new text.

\noindent Clearly, a way around those difficulties would be to try to
change the language and just remove these limitations. Scala's own
notion of traits, for example, allows state in traits. Ofcourse adding
state (and other features) to interfaces would complicate the language
and require changes in the compiler, and this would go beyond the
goals of Java 8 development team.

@@

To formalize object interfaces, we propose Featherweight Object Interfaces (FOI): 
a FeatherweightJava-style calculus, which captures the essence 
of interfaces with default methods. The semantics of object interfaces 
is given as a type-directed translation from FOI to itself. 
In the resulting FOI code all object interfaces are translated into 
regular FOI (and Java) interfaces with default methods. The translation 
is proved to be type-preserving, ensuring that the translation does 
not introduce type-errors. One interesting aspect of the theorem is that, 
if we ignore the object annotations, the program still type-checks 
under the regular typing rules of FOI (and Java).\bruno{double-check!}

We also present a prototype implementation of object interfaces using
Java annotations. One benefit of an annotation-based implementation 
is that existing Java tools (such as IDE's) work with out-of-the-box
with our implementation. As a result we could experiment object
interfaces with several interesting Java programs, and conduct various
case studies. \bruno{mention the following or not? maybe good if we
  had something about tuning.} Ofcourse, another implementation option would be to
extend Java, with a single qualifier keyword {\bf object}, so that
object interfaces would be an intrinsic part of the language. We leave
this for future work.

@@

@@

\item {\bf Object interfaces:} A simple language feature that allows 
various powerful programming idioms to be expressed conveniently in
Java. Object interfaces are pure sugar: they can be desugared into standard Java 
code, without loss of modularity or type-safety.

\item {\bf Featherweight Object (FC):} A simple formal calculus that models 
the essential features of Java 8 interfaces with default methods, and 
can be used to formally define the desugaring of object interfaces. 
We prove ...

@@

\bruno{Useful reference, perhaps, on the discussion of why not override object 
methods: 
\url{http://mail.openjdk.java.net/pipermail/lambda-dev/2013-March/008435.html}}

>>>>

\noindent Clearly, a way around those difficulties would be to try to
change the language and just remove these limitations. Scala's own
notion of traits, for example, allows state in traits. Of course adding
state (and other features) to interfaces would complicate the language
and require changes in the compiler, and this would go beyond the
goals of Java 8 development team.

@@

To formalize object interfaces, we propose Classless Java (CJ)\marco{I propose to call it directly ClassLess Java, the  Featherweight part is in being classless.}: 
a FeatherweightJava-style calculus, which captures the essence 
of interfaces with default methods. The semantics of object interfaces 
is given as a type-directed translation from CJ to itself. 
In the resulting CJ code all object interfaces are translated into 
regular CJ (and Java) interfaces with default methods. The translation 
is proved to be type-preserving, ensuring that the translation does 
not introduce type-errors. One interesting aspect of the theorem is that, 
if we ignore the object annotations, the program still type-checks 
under the regular typing rules of CJ (and Java).\bruno{double-check!}
\marco{ I do not understand what you mean. 1) annotations are part of java and CJ, so you do not need to remove them 2) programs well typed without expansion are well typed with expansion, here you seams to state the opposite}
CJ's usefulness goes beyond serving as a calculus to formalize object
interfaces. During the development process of CJ, we encountered an 
incoherent behaviour in Java's implementation of default methods that 
could be classified as a bug. \bruno{Should we report bug to the Java
  developers mailing list?}
\marco{we need to put the exact javac version}
We also present a prototype implementation of object interfaces using
Java annotations. One benefit of an annotation-based implementation 
is that existing Java tools (such as IDE's) work with out-of-the-box
with our implementation. As a result we could experiment object
interfaces with several interesting Java programs, and conduct various
case studies. 
\marco{ I removed your text here and I add after to try talking about tuning}
%I think that the qualifier object goes AGAINST the idea of tuning.
%\bruno{mention the following or not? maybe good if we
%  had something about tuning.} Of course, another implementation option would be to
%extend Java, with a single qualifier keyword {\bf object}, so that
%object interfaces would be an intrinsic part of the language. We leave
%this for future work.

@@

Moreover, we believe that our approach is an example of a more general idea, that we call \emph{Language Tuning},
a concept that sits in the middle between a lightweight language extension and a glorified library.
Thanks to Java dynamic class loading and annotation processing,  annotations can trigger type-driven transformations on classes during either loading or compilation.
There are many examples of such Language Tuning libraries,~\cite[lomboc,anotherImSearching].
Language tuning can offers many features usually implemented by a real language extension, but they do not modify the syntax; pre-existing tools can work transparently on the tuned language.

@@

\item {\bf Object Interfaces:} A simple feature that allows 
various powerful programming idioms to be expressed conveniently in
Java. Object interfaces are pure sugar: they are desugared into standard Java 
code, without loss of modularity or type-safety.

\item {\bf ClassLess Java (CJ):} A simple formal calculus that models 
the essential features of Java 8 interfaces with default methods, and 
can be used to formally define the desugaring of object interfaces. 
We prove ...

\item {\bf A Bug in the Implementation of Java 8:}\bruno{state this as
  a contrib?}\marco{yes, especially if we "suggest solutions"}

@@

\item{\bf Language Tuning} We identify this concept and we describe Project Lomboc and object interfaces as examples of Language Tuning.
\end{itemize}

\bruno{Useful reference, perhaps, on the discussion of why not override object 
methods: 
\url{http://mail.openjdk.java.net/pipermail/lambda-dev/2013-March/008435.html}}
\bruno{Maybe Marco's text can be in the conclusion, which I think is
  ok to be more provocative. I tried to be a bit more contained in the
Introduction.}
